---
title: "bi exploration"
author: "Norman Poh"
date: "1 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r}
setwd("C:/Users/npoh/Documents/Git/projects/bi/modelling")
library(palab)
library(palabmod)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)

```

## Help
```{r}
all_packages <- ls(getNamespace("palab"))
all_packages
```


## List files
```{r}
data_dir = "../data/norman_subset/"
files <- list.files(data_dir)
files
```

## Load the new predictors as they were found useful
```{r}
new_predictor_list <- read_csv("newpredictorList1.csv")
new_predictor_list$type <- as.factor(new_predictor_list$type)

```
## Generating a variable configuration file for the positive and negative samples
```{r}
var_config_generator(input_csv = paste0(data_dir, files[1]),
                     prefix = "bi_",
                     output_dir = ".")

var_config_generator(input_csv = paste0(data_dir, files[3]),
                     prefix = "bi_neg_",
                     output_dir = ".")

```
## Rectify the variable configuration file
We need to process the file bi_var_config.csv because it does not provide the correct variable type so we progmatically correct for the output in the following ways:

```{r}
config_pos <- read_csv("bi_var_config.csv")

# 1. Add key -- otherwise the file won't load
config_pos$Type[ str_detect(config_pos$Column, "PATIENT_ID") ] = "key"

# 2. Change the dates to others
### Just testing matching <- str_view(config_pos$Column,"_DT", match=TRUE)
matching <- str_detect(config_pos$Column,"_DT")
config_pos$Type[matching]="other"
write_csv(config_pos,"bi_var_config.csv")

# 3. Check the variables that are included according to the new_predictor_list
n <- length(new_predictor_list$Variable_name)
stat <- rep(0,n)

# Find if the new predictors are in the original column
is_included <- logical(length(config_pos$Column))
for(i in 1:n) {
  matching <- str_detect(config_pos$Column,new_predictor_list$Variable_name[i])
  #str_view(config_pos$Column,new_predictor_list$Variable_name[1])
  stat[i]=sum(matching)
  is_included <- is_included | matching
}
```

## It seems that their sums are not equal -- see below so there might be duplicates in the new_predictor list but we ignore this for now
```{r}
sum(is_included)
sum(stat)
```
## Filter the columns to include the variables according to the new predictor list only

```{r}
config_pos_included <- config_pos[is_included,]
View(config_pos_included)
write_csv(config_pos_included,"bi_new_predictor_var_config.csv")
```

## We process the negative samples slightly differently
```{r}
config_neg <- read_csv("bi_neg_var_config.csv")
# 1. Add key -- otherwise the file won't load
config_neg$Type[ str_detect(config_neg$Column, "NON_IPF_PATIENT") ] = "key"

# 2. Change the dates to others
### Just testing matching <- str_view(config_pos$Column,"_DT", match=TRUE)
matching <- str_detect(config_neg$Column,"_DT")
config_neg$Type[matching]="other"
write_csv(config_neg,"bi_neg_var_config.csv")
```

## Analyse the negative samples
```{r}

setdiff(config_neg$Column, config_pos$Column )
setdiff(config_pos$Column, config_neg$Column )


```
## We have the following variables and here's what we do with them 
### keep the variables (but not used as predictor)
Keep the foreign key: IPF_PATIENT

### delete -- check with Orla
IPF_PATIENT_AGE
IPF_PATIENT_GENDER
NON_IPF_INDEX_2
IDX_LATE_LOOK_BACK_DAYS

### rename -- but are we going to use these variables as predictors? (we can rename these after loading)
NON_IPF_PATIENT -> PATIENT_ID
NON_IPF_PAT_AGE -> PATIENT_AGE
NON_IPF_PAT_GENDER -> PATIENT_GENDER
NON_IPF_LOOKBACK_2 -> LOOKBACK_DAYS
NON_ASTHMA_COPD_FLAG -> ASTHMA_COPD_FLAG
NON_12_MONTHS -> IN_12_MONTHS
NON_IN_24_MONTHS -> IN_24_MONTHS
NON_IN_36_MONTHS -> IN_36_MONTHS

### Variables that are present in the positive samples but missing in the negative samples -- so these are not predictors and should be deleted -- need to confirm with Orla
 [14] "NBR_PRED_NON_PRED_SET_TO_Y" 
 [15] "LVL1_NOT_PRED_FLAG"           "LVL1_NOT_PRED_M_S_DT"        
 [17] "LVL1_NOT_PRED_CLM_CNT"        "LVL1_NOT_PRED_VST_CNT"       
 [19] "LVL1_NOT_PRED_AVG_RXDX"       "LVL2_IPF_DIAG_FLAG"          
 [21] "LVL2_IPF_DIAG_M_S_DT"         "LVL2_IPF_DIAG_CLM_CNT"       
 [23] "LVL2_IPF_DIAG_VST_CNT"        "LVL2_IPF_DIAG_AVG_RXDX"      
 [25] "LVL2_RFOPC_FLAG"              "LVL2_RFOPC_M_S_DT"           
 [27] "LVL2_RFOPC_CLM_CNT"           "LVL2_RFOPC_VST_CNT"          
 [29] "LVL2_RFOPC_AVG_RXDX"          "LVL2_IPF_RPD_KC_FLAG"        
 [31] "LVL2_IPF_RPD_KC_M_S_DT"       "LVL2_IPF_RPD_KC_CLM_CNT"     
 [33] "LVL2_IPF_RPD_KC_VST_CNT"      "LVL2_IPF_RPD_KC_AVG_RXDX"    
 [35] "LVL2_IPF_RPD_I_FLAG"          "LVL2_IPF_RPD_I_M_S_DT"       
 [37] "LVL2_IPF_RPD_I_CLM_CNT"       "LVL2_IPF_RPD_I_VST_CNT"      
 [39] "LVL2_IPF_RPD_I_AVG_RXDX"      "LVL2_OL_IPF_TRT_FLAG"        
 [41] "LVL2_OL_IPF_TRT_M_S_DT"       "LVL2_OL_IPF_TRT_CLM_CNT"     
 [43] "LVL2_OL_IPF_TRT_VST_CNT"      "LVL2_OL_IPF_TRT_AVG_RXDX"    
 [45] "LVL3_HR_CT_SC_FLAG"           "LVL3_HR_CT_SC_M_S_DT"        
 [47] "LVL3_HR_CT_SC_CLM_CNT"        "LVL3_HR_CT_SC_VST_CNT"       
 [49] "LVL3_HR_CT_SC_AVG_RXDX"       "LVL3_LUNG_BIOP_FLAG"         
 [51] "LVL3_LUNG_BIOP_M_S_DT"        "LVL3_LUNG_BIOP_CLM_CNT"      
 [53] "LVL3_LUNG_BIOP_VST_CNT"       "LVL3_LUNG_BIOP_AVG_RXDX"     
 [55] "LVL3_EH_MC_FLAG"              "LVL3_EH_MC_M_S_DT"           
 [57] "LVL3_EH_MC_CLM_CNT"           "LVL3_EH_MC_VST_CNT"          
 [59] "LVL3_EH_MC_AVG_RXDX"          "LVL3_RA_FLAG"                
 [61] "LVL3_RA_M_S_DT"               "LVL3_RA_CLM_CNT"             
 [63] "LVL3_RA_VST_CNT"              "LVL3_RA_AVG_RXDX"            
 [65] "LVL3_SICCA_FLAG"              "LVL3_SICCA_M_S_DT"           
 [67] "LVL3_SICCA_CLM_CNT"           "LVL3_SICCA_VST_CNT"          
 [69] "LVL3_SICCA_AVG_RXDX"          "LVL3_SLE_FLAG"               
 [71] "LVL3_SLE_M_S_DT"              "LVL3_SLE_CLM_CNT"            
 [73] "LVL3_SLE_VST_CNT"             "LVL3_SLE_AVG_RXDX"           
 [75] "LVL3_SYS_SCLER_FLAG"          "LVL3_SYS_SCLER_M_S_DT"       
 [77] "LVL3_SYS_SCLER_CLM_CNT"       "LVL3_SYS_SCLER_VST_CNT"      
 [79] "LVL3_SYS_SCLER_AVG_RXDX"      "LVL3_LI_SAR_FLAG"            
 [81] "LVL3_LI_SAR_M_S_DT"           "LVL3_LI_SAR_CLM_CNT"         
 [83] "LVL3_LI_SAR_VST_CNT"          "LVL3_LI_SAR_AVG_RXDX"        
 [85] "LVL3_LI_SCLER_FLAG"           "LVL3_LI_SCLER_M_S_DT"        
 [87] "LVL3_LI_SCLER_CLM_CNT"        "LVL3_LI_SCLER_VST_CNT"       
 [89] "LVL3_LI_SCLER_AVG_RXDX"       "LVL3_PI_FIBRO_FLAG"          
 [91] "LVL3_PI_FIBRO_M_S_DT"         "LVL3_PI_FIBRO_CLM_CNT"       
 [93] "LVL3_PI_FIBRO_VST_CNT"        "LVL3_PI_FIBRO_AVG_RXDX"      
 [95] "LVL3_RHEU_HD_FLAG"            "LVL3_RHEU_HD_M_S_DT"         
 [97] "LVL3_RHEU_HD_CLM_CNT"         "LVL3_RHEU_HD_VST_CNT"        
 [99] "LVL3_RHEU_HD_AVG_RXDX"        "LVL3_RHEU_PNEU_FLAG"         
[101] "LVL3_RHEU_PNEU_M_S_DT"        "LVL3_RHEU_PNEU_CLM_CNT"      
[103] "LVL3_RHEU_PNEU_VST_CNT"       "LVL3_RHEU_PNEU_AVG_RXDX"     
[105] "LVL3_RHEU_LUNG_FLAG"          "LVL3_RHEU_LUNG_M_S_DT"       
[107] "LVL3_RHEU_LUNG_CLM_CNT"       "LVL3_RHEU_LUNG_VST_CNT"      
[109] "LVL3_RHEU_LUNG_AVG_RXDX"      "LVL3_SARCOID_FLAG"           
[111] "LVL3_SARCOID_M_S_DT"          "LVL3_SARCOID_CLM_CNT"        
[113] "LVL3_SARCOID_VST_CNT"         "LVL3_SARCOID_AVG_RXDX"       
[115] "LVL3_PNEU_OH_FLAG"            "LVL3_PNEU_OH_M_S_DT"         
[117] "LVL3_PNEU_OH_CLM_CNT"         "LVL3_PNEU_OH_VST_CNT"        
[119] "LVL3_PNEU_OH_AVG_RXDX"        "LVL3_ACUT_I_PNEU_FLAG"       
[121] "LVL3_ACUT_I_PNEU_M_S_DT"      "LVL3_ACUT_I_PNEU_CLM_CNT"    
[123] "LVL3_ACUT_I_PNEU_VST_CNT"     "LVL3_ACUT_I_PNEU_AVG_RXDX"   
[125] "LVL3_CRYP_ORG_PNEU_FLAG"      "LVL3_CRYP_ORG_PNEU_M_S_DT"   
[127] "LVL3_CRYP_ORG_PNEU_CLM_CNT"   "LVL3_CRYP_ORG_PNEU_VST_CNT"  
[129] "LVL3_CRYP_ORG_PNEU_AVG_RXDX"  "LVL3_DESQ_I_PNEU_FLAG"       
[131] "LVL3_DESQ_I_PNEU_M_S_DT"      "LVL3_DESQ_I_PNEU_CLM_CNT"    
[133] "LVL3_DESQ_I_PNEU_VST_CNT"     "LVL3_DESQ_I_PNEU_AVG_RXDX"   
[135] "LVL3_ID_I_PNEU_FLAG"          "LVL3_ID_I_PNEU_M_S_DT"       
[137] "LVL3_ID_I_PNEU_CLM_CNT"       "LVL3_ID_I_PNEU_VST_CNT"      
[139] "LVL3_ID_I_PNEU_AVG_RXDX"      "LVL3_ID_I_PNEU_NOS_FLAG"     
[141] "LVL3_ID_I_PNEU_NOS_M_S_DT"    "LVL3_ID_I_PNEU_NOS_CLM_CNT"  
[143] "LVL3_ID_I_PNEU_NOS_VST_CNT"   "LVL3_ID_I_PNEU_NOS_AVG_RXDX" 
[145] "LVL3_ID_LYM_I_PNEU_FLAG"      "LVL3_ID_LYM_I_PNEU_M_S_DT"   
[147] "LVL3_ID_LYM_I_PNEU_CLM_CNT"   "LVL3_ID_LYM_I_PNEU_VST_CNT"  
[149] "LVL3_ID_LYM_I_PNEU_AVG_RXDX"  "LVL3_ID_NS_PNEU_FLAG"        
[151] "LVL3_ID_NS_PNEU_M_S_DT"       "LVL3_ID_NS_PNEU_CLM_CNT"     
[153] "LVL3_ID_NS_PNEU_VST_CNT"      "LVL3_ID_NS_PNEU_AVG_RXDX"    
[155] "LVL3_RESP_BRON_I_LD_FLAG"     "LVL3_RESP_BRON_I_LD_M_S_DT"  
[157] "LVL3_RESP_BRON_I_LD_CLM_CNT"  "LVL3_RESP_BRON_I_LD_VST_CNT" 
[159] "LVL3_RESP_BRON_I_LD_AVG_RXDX" "LVL3_AZATHIO_FLAG"           
[161] "LVL3_AZATHIO_M_S_DT"          "LVL3_AZATHIO_CLM_CNT"        
[163] "LVL3_AZATHIO_VST_CNT"         "LVL3_AZATHIO_AVG_RXDX"       
[165] "LVL3_COLCH_FLAG"              "LVL3_COLCH_M_S_DT"           
[167] "LVL3_COLCH_CLM_CNT"           "LVL3_COLCH_VST_CNT"          
[169] "LVL3_COLCH_AVG_RXDX"          "LVL3_CYCLOPH_FLAG"           
[171] "LVL3_CYCLOPH_M_S_DT"          "LVL3_CYCLOPH_CLM_CNT"        
[173] "LVL3_CYCLOPH_VST_CNT"         "LVL3_CYCLOPH_AVG_RXDX"       
[175] "LVL3_CYCLOSP_FLAG"            "LVL3_CYCLOSP_M_S_DT"         
[177] "LVL3_CYCLOSP_CLM_CNT"         "LVL3_CYCLOSP_VST_CNT"        
[179] "LVL3_CYCLOSP_AVG_RXDX"        "LVL3_MYCOPH_FLAG"            
[181] "LVL3_MYCOPH_M_S_DT"           "LVL3_MYCOPH_CLM_CNT"         
[183] "LVL3_MYCOPH_VST_CNT"          "LVL3_MYCOPH_AVG_RXDX"        
[185] "LVL3_N_ACETL_FLAG"            "LVL3_N_ACETL_M_S_DT"         
[187] "LVL3_N_ACETL_CLM_CNT"         "LVL3_N_ACETL_VST_CNT"        
[189] "LVL3_N_ACETL_AVG_RXDX"        "LVL3_SILDEN_FLAG"            
[191] "LVL3_SILDEN_M_S_DT"           "LVL3_SILDEN_CLM_CNT"         
[193] "LVL3_SILDEN_VST_CNT"          "LVL3_SILDEN_AVG_RXDX"       

## Load the negative data
```{r}
data_neg <- read_transform(
  input_csv = paste0(data_dir, files[3]), #"Non_IPF_Asthma_COPD_Cohort.csv"
  var_config_csv = "bi_neg_var_config.csv",
  read_key_as_double = TRUE
)
colnames(data_neg$data)
View(data_neg$data)
```


## Variables in the negative data set but not in the positive data set
```{r}
#setdiff(config_pos_included$Column, config_neg$Column)
additional_neg_colnames <- setdiff(config_neg$Column, config_pos_included$Column)

data_neg$data <- data_neg$data %>% rename(NON_IPF_PATIENT = PATIENT_ID)

data_neg$data <- dplyr::rename(data_neg$data, NON_IPF_PAT_AGE = PATIENT_AGE)

data_neg$data <- rename(data_neg$data, NON_IPF_PAT_AGE = PATIENT_AGE)
data_neg$data <- rename(data_neg$data, NON_IPF_PAT_GENDER = PATIENT_GENDER)
data_neg$data <- rename(data_neg$data, NON_IPF_LOOKBACK_2 = LOOKBACK_DAYS)
data_neg$data <- rename(data_neg$data, NON_ASTHMA_COPD_FLAG = ASTHMA_COPD_FLAG)
data_neg$data <- rename(data_neg$data, NON_12_MONTHS = IN_12_MONTHS)
data_neg$data <- rename(data_neg$data, NON_IN_24_MONTHS = IN_24_MONTHS)
data_neg$data <- rename(data_neg$data, NON_IN_36_MONTHS = IN_36_MONTHS)

colnames(data_neg$data)
data_neg_cleaned <- select(data_neg$data, -match( additional_neg_colnames) )

#config_neg$Column[: -c(1)]
```

## Load the positive data
```{r}
data_pos <- read_transform(
  input_csv = paste0(data_dir, files[1]),
  #var_config_csv = "bi_new_predictor_var_config.csv",
  var_config_csv = "bi_var_config.csv",
  read_key_as_double = TRUE
)
colnames(data_pos$data)
View(data_pos$data)
```
## checking
```{r}


# please input only a dataframe full of dates into this function:
create_date_diffs <- function(input, index_col = "index_date") {
  #input is the data frame
  date_cols <- input[, -which(colnames(input) == index_col)]
  
  date_diffs <- as.data.frame(sapply(date_cols, function(x) { 
    input[[index_col]] - x
    
  }))
  
  return(date_diffs)
}

datediff_ <- (dmy(data_pos$data$FIRST_IPF_DIAG_FLAG) 
  - dmy(data_pos$data$FIRST_IPF_EXPOSURE_DT)  )

data_working <- select(data_pos$data, ends_with("_DT"), FIRST_IPF_DIAG_FLAG)
data_working <- as.data.frame(  lapply(data_working, dmy) )

# data_format <- as.data.frame( 
#   lapply(data_working, function(x) { ifelse(!is.na(x), dmy(as.character( x )), NA) }) 
# ) 
# CTRL SHIFT C

# Calculate the date difference
datediff_pos <-  as.data.frame(  lapply(data_working, 
                                         function(x) {dmy(data_pos$data$FIRST_IPF_DIAG_FLAG) - x }) )

stat <- t( as_tibble( lapply(datediff_pos, function(x) {sum(is.na(x)) } ) ) )

# data_pos$data %>% mutate_if(str_detect(colnames(data_pos$data), "_DT")

ggplot(data_pos$data, aes(x=LOOKBACK_DAYS)) + geom_histogram(binwidth=.5)


```

## Univariate stats
```{r}
univar_results <- univariate_stats(input = df$data,
                                   var_config = "bi_var_config.csv",
                                   vargt0 = FALSE)
View(univar_results$cat)
```

## check
```{r}
p<-ggplot(data=univar_results$cat, aes(x=Variable, y=`Missing, prop`)) +
  geom_bar(stat="identity")
p + coord_flip()
```

## check
```{r}
p<-ggplot(data=univar_results$numerical, aes(x=Variable, y=`Missing, prop`)) +
  geom_bar(stat="identity")
p + coord_flip()
univar_results$numerical$Variable
```


## Load the negative data
```{r}
df_neg <- read_transform(
  input_csv = paste0(data_dir, files[2]),
  var_config_csv = "bi_var_config.csv",
  #missing_values = "-99, -999",
  read_key_as_double = TRUE
)
```

##