---
title: "BI-IPF with xgboost"
author: "Norman Poh"
date: "4 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r}
run_on_server <- TRUE
load_data_with_config <- TRUE

if (run_on_server) {
  setwd("K:/Norman/BI_IPF2017/modelling") 
} else {
  setwd("C:/Users/npoh/Documents/Git/projects/bi/modelling")
}

library(palab)
library(palabmod)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
library(mlr)
```

## Start with repeatability in mind
```{r results = "hide"}
random_seed <- 123
set.seed(random_seed, "L'Ecuyer")
```

## Functions

### divide_into_train_test_sets: (idlist, train_prop = .75 ) -> two idlist

```{r}
divide_into_train_test_sets <- function(id_pos, train_prop = .75 ) {
  #INPUT:
  #id_list    : a patient IDs list that is unique
  #train_prop : training proportion between 0 and 1
  #OUTPUT
  #a list with two lists of patient IDs, corresponding to the training and the test set
  
  test_prop <- 1 - train_prop
  
  mylist <- round( c(train_prop, test_prop) * length(id_pos) )
  
  trainlist <- 1:mylist[1]
  testlist <- (mylist[1]+1):length(id_pos)
  
  order <- sample( length(id_pos) )
  
  idlist <-  vector('list',2)
  idlist[[1]] <- id_pos[order[trainlist]]
  idlist[[2]] <- id_pos[order[testlist]]
  
  return(idlist)
}

```

## List files -- on local git
```{r}

if (run_on_server) {
  data_dir = ""

    files <- c(
  "K:/Hui/Project_2016/BI_IPF_2016/04_Summary/004_data/all_features_neg.csv", 
  "K:/Hui/Project_2016/BI_IPF_2016/04_Summary/004_data/all_features_pos.csv",
  "K:/Hui/Project_2016/BI_IPF_2016/04_Summary/004_data/all_features_score.csv",
  "K:/Hui/Project_2016/BI_IPF_2016/04_Summary/004_data/Ss_26m_218_features.csv")

  date_files <- c(  
  "K:/Norman/BI_IPF2017/data/raw/Non_IPF_Asthma_COPD_Cohort.csv",
  "K:/Norman/BI_IPF2017/data/raw/IPF_Cohort.csv",
  "K:/Norman/BI_IPF2017/data/raw/Representative_Sample_Asthma_COPD_Cohort.csv",
  "K:/Norman/BI_IPF2017/data/raw/Scoring_Sample_V2.csv" )

} else {
  data_dir = "../data/features/"
  files <- list.files(data_dir)
  files
  
  warning("date_files are not defined!")
}
```

## Rectify the variable configuration file
We need to process the file bi_features_var_config.csv because it does not provide the correct variable type so we programatically correct for the output in the following ways:

```{r}

if (load_data_with_config) {
  config <- read_csv("bi_features_var_config.csv")
} else {

  var_config_generator(input_csv = paste0(data_dir, files[1]),
                     prefix = "bi_features_",
                     output_dir = ".")
  
  # 1. Add key -- otherwise the file won't load
  config$Type[ str_detect(config$Column, "^patient_id$") ] = "key"
  
  # We shall not carry out step 2 below because xgboost can only take numerical data !!!
  # 2. Change the flag to categorical 
  # matching <- str_detect(config$Column,"_FLAG$")
  # config$Type[matching]="categorical"
  # 
  # matching <- str_detect(config$Column,"_BEFORE$")
  # config$Type[matching]="categorical"
  # 
  # matching <- str_detect(config$Column,"_AFTER$")
  # config$Type[matching]="categorical"
  # 
  # matching <- str_detect(config$Column,"^label$")
  # config$Type[matching]="categorical"
  # 
  write_csv(config,"bi_features_var_config.csv")
}
```

## Load three files so that they are ready for splitting

These files are neg-small, pos, neg-large, and we want to form just positive and negative samples

```{r}
bi <-  vector('list',3)
for (f in 1:3){
  bi[[f]] <- read_transform(
    input_csv = paste0(data_dir, files[f]),
    var_config_csv = "bi_features_var_config.csv",
    read_key_as_double = TRUE
  )
}
```
## The problem below is solved

<!-- --------------------------------------------------------------------------------------------- -->
<!-- ## There seems to be an error when reading the the third file so I reproduce it here -->
<!-- --------------------------------------------------------------------------------------------- -->
<!-- The error is "There are 5598822 entries but only 5598801 unique values". -->

<!-- This is documented in an email dated 6 Sept 2017 -->
<!-- ```{r} -->
<!-- f<-3 -->
<!-- ( paste0("reading ", files[f]) ) -->
<!-- bi3 <- read_transform( -->
<!--   input_csv = paste0(data_dir, files[f]), -->
<!--   var_config_csv = "bi_features_var_config.csv", -->
<!--   read_key_as_double = FALSE -->
<!--   ) -->
<!-- ``` -->

## check the patient IDs are in the matched ID set

```{r}
id_pos <- sort(unique( bi[[2]]$data$patient_id ) )
for (f in 1:3){
  mid <- sort(unique( bi[[f]]$data$matched_patient_id )) # matched id
  if (length(setdiff(id_pos, mid)) == 0 ) {
    cat(sprintf("There is no difference between positive patient id and the matched patient id for data set %d\n",f))
  } else {
    cat("The following IDs have no matched negative samples\n")
    cat( sprintf("n = %d\n", length(setdiff(id_pos, mid)) ) )
    print(setdiff(id_pos, mid))
  }
}
id_pos <- sort(unique( bi[[2]]$data$patient_id ) )
```

## Filter out this user
```{r}
for (f in 1:3){
  bi[[f]]$data <- bi[[f]]$data %>% filter(matched_patient_id != 2154663092)
}

```

## Just checking
```{r}

for (f in 1:3){
  cat ("Number of rows: ")
  cat (nrow(bi[[f]]$data) )
  cat ("\nNumber of positive samples: ")
  cat (sum(bi[[f]]$data$label))
  cat("\n")
}

```

## Divide the data into training and test sets

```{r}
if ( file.exists("idlist_partition.txt")) {
  idlist <- read_csv("idlist_partition.txt")
} else {
  idlist <- divide_into_train_test_sets(id_pos, train_prop = .75)
  write_csv()
}
data_ <-  vector('list',2)

# the training set
t <- 1
data_[[t]] <- rbind(subset( bi[[2]]$data, matched_patient_id %in% idlist[[t]]) , #pos
                    subset( bi[[1]]$data, matched_patient_id %in% idlist[[t]]) ) #neg 200
                    #subset( bi[[3]]$data, matched_patient_id %in% idlist[[t]]) ) #neg 653
# the test set
t <- 2
data_[[t]] <- rbind(subset( bi[[2]]$data, matched_patient_id %in% idlist[[t]]) , #pos
                    subset( bi[[1]]$data, matched_patient_id %in% idlist[[t]]) , #neg 200
                    subset( bi[[3]]$data, matched_patient_id %in% idlist[[t]]) ) #neg 653

# checking the prior probability of the positive class remains the same in both training and test sets
sum(data_[[1]]$label) / nrow(data_[[1]])
sum(data_[[2]]$label) / nrow(data_[[2]])

# now remove variables
# rm(bi)

```




## Load the new predictors as they were found useful
```{r}
new_predictor_list <- read_csv("newpredictorList1.csv")
new_predictor_list$type <- as.factor(new_predictor_list$type)
```

## Check which variables are included according to the new_predictor_list
```{r}
n <- length(new_predictor_list$Variable_name)
stat <- rep(0,n)

# Find if the new predictors are in the original column
is_included <- logical(length(config$Column))
for(i in 1:n) {
  matching <- str_detect(config$Column,new_predictor_list$Variable_name[i])
  #str_view(config$Column,new_predictor_list$Variable_name[1])
  stat[i]=sum(matching)
  is_included <- is_included | matching
}
```

## It seems that their sums are not equal -- see below so there might be duplicates in the new_predictor list but we ignore this for now
```{r}
sum(is_included)
sum(stat)
```

## List variables that should not have been included

Note that I'm not changing any thing here.
```{r}
var_list_to_exclude <- config$Column[is_included == FALSE]
print(var_list_to_exclude)
```

The above results look reasonable as we don't have many features that we need to ignore. So, I won't exclude anything at this stage.

patient_id, label, matched_patient_id are not needed for the modelling stage

# var_list_to_exclude <- var_list_to_exclude[c(-1,-6, -7)]

## Generate the var_config.csv files for the date_files
```{r}
load_data_with_config_raw <- FALSE

if (! load_data_with_config_raw) {

  f <- 2 # for the positive samples
  config_raw_fname <- sprintf("bi_raw%d_var_config.csv",f)
  config_raw <- read_csv(config_raw_fname)
  var_config_generator(input_csv = paste0(data_dir, date_files[f]),
                       prefix = sprintf("bi_raw%d_",f),
                       output_dir = ".")
  config_raw$Type[ str_detect(config_raw$Column, "^PATIENT_ID$") ] <- "key"
    
  # 2. Change the flag to others
  matching <- str_detect(config_raw$Column,"_DT$") | str_detect(config_raw$Column,"_DT_$")
  config_raw$Type[matching]="others"
  write_csv(config_raw, config_raw_fname)
  
  
  for (f in c(1,3)) {
    config_raw_fname <- sprintf("bi_raw%d_var_config.csv",f)
    
    var_config_generator(input_csv = paste0(data_dir, date_files[f]),
                         prefix = sprintf("bi_raw%d_",f),
                         output_dir = ".")
    config_raw <- read_csv(config_raw_fname)
    matching <-  str_detect(config_raw$Column, "^NON_IPF_PATIENT$")
    config_raw$Type[ matching ] <- "key"
    config_raw$Type[ ! matching ] <- "others"  
    
    write_csv(config_raw, config_raw_fname)
  }

}
```

## Now load the date files

```{r}
#bi_raw <-  vector('list',3)
for (f in c(1, 3) ){
  config_raw_fname <- sprintf("bi_raw%d_var_config.csv",f)
  bi_raw[[f]] <- read_transform(
    input_csv = paste0(data_dir, date_files[f]),
    var_config_csv = config_raw_fname,
    read_key_as_double = TRUE
  )
}

for (f in 1:3){
 print( nrow(bi_raw[[f]]$data)  )
}
```

## Rename the column

```{r}
bi_raw[[1]]$data <- rename(bi_raw[[1]]$data, PATIENT_ID = NON_IPF_PATIENT)
bi_raw[[3]]$data <- rename(bi_raw[[3]]$data, PATIENT_ID = NON_IPF_PATIENT)
```

## below wound not have been necessary now
<!-- ## Rectify my error :( -->
<!-- ```{r} -->

<!-- for (f in 1:3){ -->
<!--   print( nrow(bi[[f]]$data)) -->
<!-- } -->
<!-- for (f in 1:3){ -->
<!--   print( nrow(bi_raw[[f]]$data)) -->
<!-- } -->

<!-- # rectify the problem -->
<!-- bi_raw_ <- vector('list', 3) -->
<!-- bi_raw_[[1]] <- bi_raw[[2]] -->
<!-- bi_raw_[[2]] <- bi_raw[[1]] -->
<!-- bi_raw_[[3]] <- bi_raw[[3]] -->

<!-- rm(bi_raw) -->
<!-- bi_raw <- bi_raw_ -->
<!-- rm(bi_raw_) -->

<!-- for (f in 1:3){ -->
<!--   print( nrow(bi[[f]]$data)) -->
<!-- } -->
<!-- for (f in 1:3){ -->
<!--   print( nrow(bi_raw[[f]]$data)) -->
<!-- } -->
<!-- ``` -->

## Align bi and bi_raw by retaining only what is available in bi, discarding extra rows in bi_raw 
This is to ensure that only rows where the matching patient ids (matched_patient_id) for the patient_id are included

```{r}
for (f in 1:3){
  key_ <- bi[[f]]$data$patient_id
  print( length(key_) )
  print( length(bi_raw[[f]]$data$PATIENT_ID) )
  #nrow ( subset( bi_raw[[f]]$data, PATIENT_ID %in% key_) )
  bi_raw[[f]] <- subset( bi_raw[[f]]$data, PATIENT_ID %in% key_)
}
```

## Checking

```{r}
for (f in 1:3){
  key_ <- bi[[f]]$data$patient_id
  print( nrow(bi_raw[[f]]) )
  print( length(key_) )
  
}
```

## Save the bi and bi_raw variables which are now completely aligned

```{r}
saveRDS(bi_raw, "bi_raw.rds")
saveRDS(bi, "bi.rds")
```

<!-- ##  -->

<!-- ```{r} -->
<!-- id_pos <- sort(unique( bi[[2]]$data$patient_id ) ) -->

<!-- idlist <- divide_into_train_test_sets(id_pos, train_prop = .75) -->
<!-- ``` -->

## Let's derive new variables


```{r}
config_raw <- vector('list', 3)
for (f in 1:3) {
  config_raw_fname <- sprintf("bi_raw%d_var_config.csv",f)
  config_raw[[f]] <- read_csv(config_raw_fname)
}
```

## Find out common_fields_DT, i.e., the common fields across the labeled data sets ending with a date

```{r}
# First there should be no difference between sets 1 and 3 (both neg samples)
setdiff(config_raw[[3]]$Column, config_raw[[1]]$Column)
setdiff(config_raw[[1]]$Column, config_raw[[3]]$Column)

pos_fields <- as_tibble( setdiff(config_raw[[2]]$Column, config_raw[[1]]$Column) )
common_fields <- as_tibble( intersect( config_raw[[2]]$Column, config_raw[[1]]$Column) )

pos_fields_only <- setdiff(pos_fields, common_fields)
View(pos_fields_only)
View(common_fields)
common_fields_DT <- common_fields %>% filter( str_detect(value, "_DT$")  )

```

## positive set has three possible fields

```{r}
diagnosis <- c('FIRST_OFEV_RX_DT', 'FIRST_ESBRIET_RX_DT', 'FIRST_IPF_DIAG_FLAG')
#bi_raw[[f]]$

```
FIRST_IPF_EXPOSURE_DT might be derived from the above

## negative set has three possible fields
NON_IPF_INDEX_2
IDX_LATE_LOOK_BACK_DAYS



```{r}
data_working <- select(bi_raw, ends_with("_DT"), FIRST_IPF_DIAG_FLAG)
data_working <- as.data.frame(  lapply(data_working, dmy) )

# Calculate the date difference
datediff_pos <-  as.data.frame(  lapply(data_working, 
                                         function(x) {dmy(data_pos$data$FIRST_IPF_DIAG_FLAG) - x }) )


```






